Efficient, Robust and Exact Boolean Operations on N Triangular Mesh Primitives

### 2016-10-15 summary.tex
10.15

During face classification, we evaluate whether a given face $s$ belongs to the final geometry according to the n-primitive boolean function $f$ :
\begin{equation}
\lambda_f(s) = f(\boldsymbol{\Lambda}(s)) = f(\lambda_1(s), \lambda_2(s), \cdots, \lambda_n(s)).
\end{equation}
The parameter $\lambda_i(s)$ is space indicator with respect to mesh $P_i$. To compute $\lambda_f(s)$ and classify $s$, one has to know the space indicators with respect to all the primitives (denoted as $\boldsymbol{\Lambda}(s)$). A space indicator has four conditions: completely inside (\emph{in}), completely outside (\emph{out}), on the boundary with consistent normals (\emph{same}) or opposite normals (\emph{oppo}). The rules of boolean function evaluation are described in \cite{douze2015quickcsg,feito2013fast}.

\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{falseclass}
\caption{In a 2D view, face $s$ is on the surface of mesh $P$. However, because we use barycenter $v_c(s)$ to compute the indicator  and the coordinates of $v_c(s)$ contain round-off error, we might falsely classify $s$ as outside of $P$.}
\label{fig:falseclass}
\end{figure}

@@

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=1.7in]{sgen}\\
  \caption{Triangle-triangle intersections made by different triangle pairs may (a) overlap or (b) intersect with each other. (c) Such situation introduces extra vertex which is intersection of triangles from three different meshes.}\label{fig:p-reps}
\end{figure}

@@

Using Plane-based representation, each face $s$ with $n$ edges is represented by a supporting plane $\bm{p}_{s,sp}$, where the face lies, and a list bounding planes $\{\bm{p}_{s,b}^i \ \vert\  i = 0, 1,...,n-1\}$. Each edge line $\bm{e}_s^i$ is represented by intersection $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i$. Corner vertex $\bm{v}_s^i$ is represented by intersection $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i \cap \bm{p}_{s,b}^{{(i+1)}\bmod{n}}$.

@@

\begin{figure}[t]
\centering
\includegraphics[width=2in]{iisect}
\caption{suppose? actual?}
\label{fig:iisect}
\end{figure}

@@

The intersection vertices are not only introduced by intersection between an edge and a triangle, but also introduced by intersections of three triangles. It is not enough to find all of them by only triangle-triangle intersection tests. Also, because of degenerate cases, there will be other forms of intersections involving three or even more triangles. To make tessellation easier and more robust, we have to refine these intersections and all unusual conditions have to be removed before it.

Intersection refinement is performed in a local scope. For each intersecting face $t_i$, we collect all intersections on $t$ as a set $\bm{\Gamma}(t)=\{\bm{\mathcal{I}}_{t}^i, i=0,1,2,...\}\cup\{\bm{\mathcal{I}}_{t}^{e_i},i=0,1,2\}$ and refine them together. The set $\{\bm{\mathcal{I}}_{t}^{e_i},i=0,1,2\}$ is the PBI-reps of the three edges of $t$. We treat edges as intersections for simplicity in this stage. The neighborhood component $\bm{\Psi}(\bm{\mathcal{I}}_{t}^{e_i})$ is set as $\bm{0}$. Other components of $\bm{\mathcal{I}}_{t}^{e_i}$ are straightforward to determine. Refinement on $\bm{\Gamma}(t)$ is done using only plane-based geometry predicates by the following three steps:

\vspace{0.5em}
\noindent \textbf{Collinear splitting}~~~~
Each intersection $\bm{\mathcal{I}}_{t}^x$ are testing whether they are collinearly overlapping with other intersections. If yes, $\bm{\mathcal{I}}_{t}^x$ is split by the end points of overlapping intersections (Fig. x). Because $\bm{\mathcal{I}}_{t}^x$ may collinearly overlap more than one other intersection, the splitting may be deferred until all splitting points are found. The subroutine linear order of points (\S\ref{sec:substrates}) is used to sort these splitting points along $\bm{\mathcal{I}}_{t}^x$. The PBI-reps of split segments inherit their father's except the second component the end points.

\vspace{0.5em}
\noindent \textbf{Coincidence elimination}~~~~
We merge coincidence intersections which have the same end points. Intersections are undirected so intersections with inverse end points are also coincident. The PBI-rep of the new intersection is inherited from either of the old one's except the neighborhood component $\bm{\Psi}$. We follow the merging rules below:
\begin{equation}
  \Psi_k(\bm{\mathcal{I}}_{t}^{12})=\left\{
   \begin{array}{l}
   \Psi_k(\bm{\mathcal{I}}_{t}^1)\mbox{ or }\Psi_k(\bm{\mathcal{I}}_{t}^2),\ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^1)=\Psi_k(\bm{\mathcal{I}}_{t}^2) \\
   \Psi_k(\bm{\mathcal{I}}_{t}^1),\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^2)=0  \\
   \Psi_k(\bm{\mathcal{I}}_{t}^2),\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^1)=0  \\
   \end{array}
  \right.
\end{equation}
, where $\bm{\mathcal{I}}_{t}^1$ and $\bm{\mathcal{I}}_{t}^2$ are old intersections and $\bm{\mathcal{I}}_{t}^{12}$ is the new one.

\vspace{0.5em}
\noindent \textbf{Crossing intersection resolving}~~~~
After the process of two previous steps, there are no overlapping intersections in $\bm{\Gamma}(t)$. We check whether any pair of intersections intersect each other. The definition of \emph{intersect} does not include the situation that two intersections share end points. In Fig. x, we illustrate the only two situations. In both situations, one intersection is split by other, and the same splitting routine is used as in collinear splitting.

@@

\label{sec:classification}
We traverse each face in the hybrid meshes, and classify which one belongs to the final mesh. Classification is done by evaluating the indicator vector of each face. The basic idea of our classification method is to utilize the space coherence of face indicator vectors to reuse the classification results. Also, because we store the PBI-reps which contains neighborhood information in the edges which intersect other meshes, the search space of point-in-polyhedron test is limited to only a few faces, which makes the indicator computation very fast. Also, by using P-reps, our classification method can deal with any degenerate cases.

@@

The space coherence of indicator vectors means neighboring faces could share the same indicator vector, or most components of indicator vectors. Therefore, we start from a seed face $s_0$ with its indicator vector $\bm{\Lambda}(s_0)$ known, and propagate to other faces in a flood-filling manner. The trace of indicator propagation is:
\begin{equation}
\label{eq:trace}
s_0\to s_1\to s_2\to s_3\to ...
\end{equation}

Our hybrid meshes facilitate this process by accelerating the indicator propagation. Given adjacent faces $s_1$ and $s_2$, it is straightforward to know whether $\bm{\Lambda}(s_1)$ and $\bm{\Lambda}(s_2)$ are different. The two indicator vectors are the same unless the shared edge $e_{12}$ lies on the boundary of some meshes, and it means there is an intersection $\bm{\mathcal{I}}(e_{12})$ stored on $e_{12}$. For a certain indicator $\lambda_k$, it always stands:
\begin{equation}
\label{eq:eq1}
{\lambda_k}(s_1) = {\lambda_k}(s_2),\ \ \  \mbox{if }\Psi_k(\bm{\mathcal{I}}(e_{12}))= 0.
\end{equation}
Even if $\Psi_k(\bm{\mathcal{I}}(e_{12})) \neq 0$, both $\lambda_k(t_1)$ and $\lambda_k(t_2)$ can be computed efficiently and exactly according to $\bm{\mathcal{I}}(e_{12})$ (\S\ref{sec:propagation}). We outline our classification method in Algorithm \ref{code:floodfill}.

\begin{algorithm}
\caption{Fast Face Classification}
\label{code:floodfill}
\textbf{Input: } IFH meshes $\bm{P'}$, a boolean function $f$

\textbf{Output: } Classification $f(\bm{\Lambda}(s_i))$, for all faces $s_i$ from $\bm{P'}$

\begin{algorithmic}[1]
\State Select a proper seed face $s_0$
\State Compute the first indicator vector $\boldsymbol{\Lambda}(s_0)$
\State \Call{propagate} { $s_0$ , $\boldsymbol{\Lambda}(s_0)$}
\State
\Function{propagate}{ $s$ , $\boldsymbol{\Lambda}(s)$}
    \State Compute $f(\boldsymbol{\Lambda}(s))$
    \For {each neighboring face $s_i$}
        \If {$s_i$ has been classified}
            \State continue
        \EndIf
        \If {there are PBI-reps $\{{\bm{\mathcal{I}}}_k\}$ on $e(s_i, s)$}
            \State compute $\boldsymbol{\Lambda}(s_i)$ by $\boldsymbol{\Lambda}(s)$ and $\{{\bm{\mathcal{I}}}_k\}$
            \State \Call{propagate} { $s_i$ , $\boldsymbol{\Lambda}(s_i)$}
        \Else
            \State \Call{propagate} { $s_i$ , $\boldsymbol{\Lambda}(s)$}
        \EndIf
    \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

However, using point on the boundary instead of an inner one has another problem. Given a face $s$ (may not be triangle) and one of its vertex $v_b^i(s)$, the vertex indicator $\boldsymbol{\Lambda}(v_{i}(s), P_k)$ may cause ambiguity to deduce the face indicator $\boldsymbol{\Lambda}(s, P_k)$. In {\color{red}{Fig. ?}}, $\boldsymbol{\Lambda}(v_b^1(s), P_k) = on$, then $\boldsymbol{\Lambda}(s, P_k)$ can be any of the four conditions. We solve this problem by trying other vertices in $s$ for ambiguous indicators. If all vertices indicators are $on$ and the ambiguity cannot be resolved, the current face is not proper for being the seed and we select a new one instead.

@@

\subsection{Indicator Propagation}
\label{sec:propagation}
Now, our meshes consist of vertices, edges and faces. Classification is required to perform on face level. However, polyhedron in-out test is performed based on points. Conventionally, face inner point such as barycenter is used to compute the indicator of the whole face. This is because in intersection-free meshes, face is classified as a whole, and every inner point will have the same indicator as the face's. However, the coordinates of inner points cannot be exactly represented by float point numbers, which may lead to wrong predicates, especially when the face lies on the surface of other mesh (Fig. x). Therefore, we use the vertices, whose exact representations are known, for indicator computation.

%For example, given a face $s$ and one of its vertex $v^i(s)$, if $\lambda_k(v^i(s))=on$, $\lambda_k(s)$ can be any of $on$, $in$ or $out$ (see {\color{red}{Fig. ?}})
However, there are two gaps between face indicator and vertex indicator. First, the indicator of vertex does not always equal to indicator of face. Second, face indicator has two conditions in $on$ case, $same$ and $oppo$, because the face has its orientation.

To solve these problems, we discuss our indicator trace (\ref{eq:trace}) in finer granularity. In the beginning, we start from a seed vertex $v_0$ on seed face $s_0$. We assume indicator vector $\bm{\Lambda}(v_0)$ is known. Then we use $\bm{\Lambda}(v_0)$ to compute $\bm{\Lambda}(s_0)$. In addition, we add the edge layer into the trace. Then it changes to:
\begin{equation}
v_0\to e_0\to s_0\to e_1\to s_1\to e_2\to s_2\to ...
\end{equation}
, where $v, e$ means vertex and edge respectively. We find that there are three kinds of basic operations in the trace: $v_e\to e$, $e_s\to s$ and $s\to e_s$, where $v_e$ means the vertex is the end point of $e$, and $e_s$ means the edge is the edge of $s$.

Given the partial order $on \succ in$ and $on \succ out$, our key observation is that the following relation always stands for indicators:
\begin{equation}
\label{eq:porder}
\lambda_k(v_{e_s}) \succeq \lambda_k(e_s) \succeq \lambda_k(s)
\end{equation}
, where $\lambda_k(x)$ means the indicator of $x$ against a certain primitive $P_k$. This relation can be inferred from continuous space assumption and the definition of intersection-free mesh (def. \ref{eq:ifmdef}). It tells us when we perform $v_e\to e$ and $e_s\to s$, we decide whether the $on$ indicator change to $in$ or $out$ or remain $on$. When we perform $s\to e_s$, we decide whether $in$ and $out$ indicators change to $on$. In addition, considering the orientation of face, we also need to figure out whether the $on$ indicators change to $same$ or $oppo$ during $e_s\to s$ operation.

\vspace{0.5em}
\noindent\textbf{$\bm{s\to e_s}$}~~~~ When $\lambda_k(e_s)=on$, there will be PBI-rep $\bm{\mathcal{I}}(e_s)$ attached to $e_s$ and the related neighborhood $\Psi_k(\bm{\mathcal{I}}(e_s))$ must be nonzero, since it is on the surface of mesh $P_k$:
\begin{equation}
\label{eq:eq1}
\lambda_k(e_s)=on,\ \ \  \mbox{iff. }\Psi_k(\bm{\mathcal{I}}(e_s))\neq 0.
\end{equation}
So all we need is to iterate over all nonzero components of $\bm{\Psi}(\bm{\mathcal{I}}(e_s))$ and change corresponding indicators to $on$.

\vspace{0.5em}
\noindent\textbf{$\bm{e_s\to s}$}~~~~From eq. \ref{eq:porder}, we know if $\lambda_k(e_s) \neq on$, then $\lambda_k(e_s)=\lambda_k(s)$. When $\lambda_k(e_s) = on$, we figure out $\lambda_k(s)$ using $\Psi_k(\bm{\mathcal{I}}(e_s))$. From eq. \ref{eq:eq1}, we know $\Psi_k(\bm{\mathcal{I}}(e_s)) \neq 0$ and can be either a triangle face or an edge. If $\Psi_k(\bm{\mathcal{I}}(e_s))$ is a face, we can build a trivial BSP \cite{thibault1987set} using the face on the space near $e_s$. If $\Psi_k(\bm{\mathcal{I}}(e_s))$ is an edge, we can still build BSP using the opposite two faces along $e_s$. In both cases, the BSP can be used to compute $\lambda_k(s)$ if we can sample a point from $s$ that near $e_s$. Unfortunately, we cannot guarantee to find such a point with precise coordinates. Instead, we choose a point $v_x$ from the half plane $\bm{p}_{s, sp}^+$, which is on the supporting plane of $s$ and on the positive side of bounding plane $\bm{p}_{s, b}^{e_s}$ (Fig. x). For polygons, there is at least one such $v_x$ among the vertices of $s$. We find that the BSP is so simple that indicators $\lambda_k$ of all points on $\bm{p}_{s, sp}^+$ are the same. Since $\bm{p}_{s, sp}^+ \cap s \neq \varnothing$, we can compute $\lambda_k(s)$ by figuring out $\lambda_k(v_x)$. Also, if we assign each BSP splitting plane a normal by the normal of triangle it contains, we can use it to decide the orientation ($same$ or $oppo$) when $v_x$ is on a splitting plane, which means $\lambda_k(s)=on$.

\vspace{0.5em}
\noindent\textbf{$\bm{v_e\to e}$}~~~~ While this is yet another propagation from low dimension to high dimension, there is no neighborhood information stored on vertex. If $\lambda_k(v_e)=on$, we first check whether $\lambda_k(e) = on$ by (\ref{eq:eq1}). If not, we need to find where $v_e$ lies on the surface of $P_k$ and determine the neighborhood of $v_e$. In most cases, it is straightforward since coincident vertices among different intersection-free meshes are shared. Exceptions are discussion is in \S\ref{sec:esubroutine}. The neighborhood may be a vertex, an edge or a face. After computing the neighborhood, the same BSP-based classification method is used as in ${e_s\to s}$ to determine the high dimensional $\lambda_k(e)$. We use the other vertex of $e$ to be the sample point $v_x$ for the BSP classification.

@@

Given a face $t_i$ to be classified and an I-rep $\boldsymbol\Gamma$ of intersection on one of $t_i$'s edge, we denote the three corners it as $v_{b*}^{0}(t_i), v_{b*}^{1}(t_i), v_b^2(t_i),$ where the subscript $*$ means the vertices are on the edge where $\boldsymbol\Gamma$ lies. We also denote the context component of $\boldsymbol\Gamma$ as $C_{P_h \backslash P_j}$, meaning $t_i$ is from primitive $P_j$ and intersects primitive $P_h$. From previous discussion (section \ref{sec:ir}){\color{red}{may be more than one}}, we know there are two types of intersection context---face neighborhood and edge neighborhood. In both conditions, the one or two faces involving in the context split the space into two in and out space by its (their) orientation(s). We will see that under such division (denoted as $\boldsymbol{D}(C_{P_h \backslash P_j})$), we can compute $\lambda(t_i, P_j)$ exactly.

>>>>

During face classification, we evaluate whether a given face $\bm{s}$ belongs to the final geometry according to the n-primitive boolean function $f$ :
\begin{equation}
\lambda_f(s) = f(\boldsymbol{\Lambda}(s)) = f(\lambda_1(s), \lambda_2(s), \cdots, \lambda_n(s)).
\end{equation}
The parameter $\lambda_i(s)$ is space indicator with respect to mesh $P_i$. To compute $\lambda_f(s)$ and classify $\bm{s}$, one has to know the space indicators with respect to all the primitives (denoted as $\boldsymbol{\Lambda}(s)$). A space indicator has four conditions: completely inside (\emph{in}), completely outside (\emph{out}), on the boundary with consistent normals (\emph{same}) or opposite normals (\emph{oppo}). The rules of boolean function evaluation are described in \cite{douze2015quickcsg,feito2013fast}.

\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{falseclass}
\caption{In a 2D view, face $\bm{s}$ is on the surface of mesh $P$. However, because we use barycenter $v_c(s)$ to compute the indicator  and the coordinates of $v_c(s)$ contain round-off error, we might falsely classify $\bm{s}$ as outside of $P$.}
\label{fig:falseclass}
\end{figure}

@@

\iffalse
\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=1.7in]{sgen}\\
  \caption{Triangle-triangle intersections made by different triangle pairs may (a) overlap or (b) intersect with each other. (c) Such situation introduces extra vertex which is intersection of triangles from three different meshes.}\label{fig:p-reps}
\end{figure}
\fi

@@

Using Plane-based representation, each face $\bm{s}$ with $n$ edges is represented by a supporting plane $\bm{p}_{s,sp}$, where the face lies, and a list bounding planes $\{\bm{p}_{s,b}^i \ \vert\  i = 0, 1,...,n-1\}$. Each edge line $\bm{e}_s^i$ is represented by intersection $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i$. Corner vertex $\bm{v}_s^i$ is represented by intersection $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i \cap \bm{p}_{s,b}^{{(i+1)}\bmod{n}}$.

@@

\begin{figure}[t]
\centering
\includegraphics[width=2in]{iisect}
\caption{suppose? actual?plane-triples}
\label{fig:iisect}
\end{figure}

@@

The intersection vertices are not only introduced by intersection between an edge and a triangle, but also introduced by intersections of three triangles. In order to find all intersection vertices, it is not enough by only triangle-triangle intersection tests. To make tessellation easier and more robust, we have to refine these intersections and resolve the intersections between intersections.

Intersection refinement is performed in a local scope. For each intersecting face $t$ from mesh $P_t$, we collect all intersections on $t$ into a set $\bm{\Gamma}(t)=\{\bm{\mathcal{I}}_{t}^i, i=0,1,2,...\}\cup\{\bm{\mathcal{I}}_{t}^{\bm{e}_i},i=0,1,2\}$ and refine them together. The set $\{\bm{\mathcal{I}}_{t}^{\bm{e}_i},i=0,1,2\}$ is the PBI-reps of the three edges of $t$. We treat edges as special intersections for simplicity, since edges also participate in tessellation. The neighborhood components of the edge $\bm{\Psi}(\bm{\mathcal{I}}_{t}^{\bm{e}_i})$ is set as $\{(P_t, \bm{e}_i )\}$. Other components of $\bm{\mathcal{I}}_{t}^{\bm{e}_i}$ can be determined by the P-reps of $\bm{e}_i$. The refinement on $\bm{\Gamma}(t)$ is done using only plane-based geometry predicates by the following three steps (Fig. x):

\vspace{0.5em}
\noindent \textbf{Collinear splitting}~~~~
Each intersection $\bm{\mathcal{I}}_{t}^x$ are testing whether they are collinearly overlapping with other intersections. If yes, $\bm{\mathcal{I}}_{t}^x$ is split by the end points of overlapping intersections (Fig. x). Because $\bm{\mathcal{I}}_{t}^x$ may collinearly overlap more than one other intersection, the splitting may be deferred until all splitting points are found. The subroutine linear order of points (\S\ref{sec:substrates}) is used to sort these splitting points along $\bm{\mathcal{I}}_{t}^x$. The PBI-reps of split segments inherit their father's except the third component which identifies the end points.

\vspace{0.5em}
\noindent \textbf{Coincidence elimination}~~~~
We merge coincidence intersections which have the same end points. Intersections are undirected so intersections with inverse end points are also coincident. The PBI-rep of the new intersection is inherited from either of the old one's except the neighborhood component $\bm{\Psi}$. The value element of the new neighborhood is the union of the two old ones'.
\iffalse 
 We follow the merging rules below:
\begin{equation}
  \Psi_k(\bm{\mathcal{I}}_{t}^{12})=\left\{
   \begin{array}{l}
   \Psi_k(\bm{\mathcal{I}}_{t}^1)\mbox{ or }\Psi_k(\bm{\mathcal{I}}_{t}^2),\ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^1)=\Psi_k(\bm{\mathcal{I}}_{t}^2) \\
   \Psi_k(\bm{\mathcal{I}}_{t}^1),\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^2)=0  \\
   \Psi_k(\bm{\mathcal{I}}_{t}^2),\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^1)=0  \\
   \end{array}
  \right.
\end{equation}

, where $\bm{\mathcal{I}}_{t}^1$ and $\bm{\mathcal{I}}_{t}^2$ are old intersections and $\bm{\mathcal{I}}_{t}^{12}$ is the new one.
\fi

\vspace{0.5em}
\noindent \textbf{Crossing intersection resolving}~~~~
After the process of two previous steps, there are no overlapping intersections in $\bm{\Gamma}(t)$. We check whether any pair of intersections intersect each other. The definition of \emph{intersect} does not include the situation that two intersections share end points. In Fig. x, we illustrate the two valid situations. In both situations, one intersection is split by the other, and the same splitting routine is used as in collinear splitting.

@@

\label{sec:classification}
We traverse each face in the intersection-free meshes, and classify which one belongs to the final mesh. Classification is done by evaluating the indicator vector of each face. The basic idea of our classification method is to utilize the space coherence of face indicator vectors to reuse the classification results. Because we store the PBI-reps which contains neighborhood information in the edges which intersect other meshes, the search space of point-in-polyhedron test is limited to only a few faces, which makes the indicator computation very fast. Also, by using P-reps, our classification method can deal with any degenerate cases.

@@

The space coherence of indicator vectors means neighboring faces could share the same indicator vector, or most components of indicator vectors. Therefore, we start from a seed face $\bm{s}_0$ with its indicator vector $\bm{\Lambda}(\bm{s}_0)$ known, and propagate to other faces in a flood-filling manner. The trace of indicator propagation is:
\begin{equation}
\label{eq:trace}
\bm{s}_0\to \bm{s}_1\to \bm{s}_2\to \bm{s}_3\to ...
\end{equation}

Our hybrid meshes facilitate this process by accelerating the indicator propagation. Given adjacent faces $\bm{s}_1$ and $\bm{s}_2$, it is straightforward to know whether $\bm{\Lambda}(\bm{s}_1)$ and $\bm{\Lambda}(\bm{s}_2)$ are different. The two indicator vectors are the same unless the shared edge $\bm{e}_{12}$ lies on the boundary of some meshes, and it means there is an intersection (denoted as $\bm{\mathcal{I}}(\bm{e}_{12})$) stored on $\bm{e}_{12}$. The neighborhood of $\bm{\mathcal{I}}(\bm{e}_{12})$ indicates which components of indicator vector differ between $\bm{s}_1$ and $\bm{s}_2$.
 \iffalse
 Namely, for a certain indicator $\lambda_k$ of $\bm{s}_1$ and $\bm{s}_2$, it always stands:
\begin{equation}
\label{eq:eq1}
{\lambda_k}(\bm{s}_1) = {\lambda_k}(\bm{s}_2),\ \ \  \mbox{if }\Psi_k(\bm{\mathcal{I}}(e_{12}))= 0.
\end{equation}
\fi
If there is a neighborhood $(P_k, \Psi_k) \in \bm{\Psi}(\bm{\mathcal{I}}(\bm{e}_{12}))$, both $\lambda_k(\bm{s}_1)$ and $\lambda_k(\bm{s}_2)$ can be computed efficiently and exactly according to $\Psi_k$ (\S\ref{sec:propagation}). We outline our classification method in Algorithm \ref{code:floodfill}.

\begin{algorithm}
\caption{Fast Face Classification}
\label{code:floodfill}
\textbf{Input: } Intersection-free hybrid meshes, boolean function $f$

\textbf{Output: } Classification $f(\bm{\Lambda}(\bm{s}_i))$ for all faces $\bm{s}_i$

\begin{algorithmic}[1]
\State Select a proper seed face $\bm{s}_0$
\State Compute the seed indicator vector $\boldsymbol{\Lambda}(\bm{s}_0)$
\State \Call{propagate} { $\bm{s}_0$ , $\boldsymbol{\Lambda}(\bm{s}_0)$}
\State
\Function{propagate}{ $\bm{s}$ , $\boldsymbol{\Lambda}(s)$}
    \State Compute $f(\boldsymbol{\Lambda}(s))$
    \For {each neighboring face $\bm{s}_{s, i}$}
        \If {$\bm{s}_i$ has been classified}
            \State continue
        \EndIf
        \If {there are PBI-reps ${\bm{\mathcal{I}}}_k$ on $\bm{e}(\bm{s}_{s, i}, s)$}
            \State compute $\boldsymbol{\Lambda}(\bm{s}_{s, i})$ by $\boldsymbol{\Lambda}(s)$ and ${\bm{\mathcal{I}}}_k$
            \State \Call{propagate} { $\bm{s}_{s, i}$ , $\boldsymbol{\Lambda}(\bm{s}_{s, i})$}
        \Else
            \State \Call{propagate} { $\bm{s}_{s, i}$ , $\boldsymbol{\Lambda}(s)$}
        \EndIf
    \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

However, using point on the boundary instead of an inner one has another problem. Given a face $\bm{s}$ (may not be triangle) and one of its vertex $v_b^i(s)$, the vertex indicator $\boldsymbol{\Lambda}(v_{i}(s), P_k)$ may cause ambiguity to deduce the face indicator $\boldsymbol{\Lambda}(s, P_k)$. In {\color{red}{Fig. ?}}, $\boldsymbol{\Lambda}(v_b^1(s), P_k) = on$, then $\boldsymbol{\Lambda}(s, P_k)$ can be any of the four conditions. We solve this problem by trying other vertices in $\bm{s}$ for ambiguous indicators. If all vertices indicators are $on$ and the ambiguity cannot be resolved, the current face is not proper for being the seed and we select a new one instead.

@@

\subsection{Indicator Propagation}
\label{sec:propagation}
Meshes consist of vertices, edges and faces. Classification is required to perform on the face level. However, polyhedron in-out test is performed based on points. Gaps exist between the them. Conventionally, face barycenter is used to compute the indicator of the whole face. This is because in intersection-free meshes, face is classified as a whole, and every face inner point will have the same indicator as the face's. However, the computation of inner point coordinates requires geometry construction, which introduces errors without exact arithmetic. Therefore, we use the vertices instead, whose exact representations are known, for indicator computation.

%For example, given a face $\bm{s}$ and one of its vertex $v^i(s)$, if $\lambda_k(v^i(s))=on$, $\lambda_k(s)$ can be any of $on$, $in$ or $out$ (see {\color{red}{Fig. ?}})
However, there are two gaps between face indicator and vertex indicator. First, the indicator of vertex does not always equal to indicator of face. Second, face indicator has two conditions in $on$ case---$same$ and $oppo$, because the face has its orientation.

To overcome the gaps, we discuss our indicator trace (\ref{eq:trace}) in finer granularity. In the beginning, we start from a seed vertex $\bm{v}_0$ on seed face $\bm{s}_0$. We assume indicator vector $\bm{\Lambda}(\bm{v}_0)$ is known. Then we use $\bm{\Lambda}(\bm{v}_0)$ to compute $\bm{\Lambda}(\bm{s}_0)$. In addition, we add the edge layer into the trace. Then it changes to:
\begin{equation}
\bm{v}_0\to \bm{e}_0\to \bm{s}_0\to \bm{e}_1\to \bm{s}_1\to \bm{e}_2\to \bm{s}_2\to ...
\end{equation}
, where $\bm{v}, \bm{e}$ means vertex and edge respectively. We find that there are three kinds of basic operations in the trace: $\bm{v}_{\bm{e}}\to \bm{e}$, $\bm{e}_s\to \bm{s}$ and $\bm{s}\to \bm{e}_s$, where $\bm{v}_{\bm{e}}$ means the end point of $\bm{e}$, and $\bm{e}_s$ means the edge of $\bm{s}$.

Given the partial order $on \succ in$ and $on \succ out$, our key observation is that the following relation always stands for indicators:
\begin{equation}
\label{eq:porder}
\lambda_k(\bm{v}_{\bm{e}_s}) \succeq \lambda_k(\bm{e}_s) \succeq \lambda_k(s)
\end{equation}
, where $\lambda_k(x)$ means the indicator of $x$ against a certain primitive $P_k$. This relation can be inferred from continuous space assumption and the definition of intersection-free mesh. It tells us when we perform $\bm{v}_{\bm{e}}\to \bm{e}$ and $\bm{e}_{\bm{s}}\to \bm{s}$, we decide whether the $on$ indicator change to $in$ or $out$ or remain $on$. When we perform $\bm{s}\to \bm{e}_s$, we decide whether $in$ and $out$ indicators change to $on$. In addition, considering the orientation of face, we also need to figure out whether the $on$ indicators change to $same$ or $oppo$ during $\bm{e}_s\to s$ operation.

\vspace{0.5em}
\noindent\textbf{$\bm{s\to \bm{e}_s}$}~~~~ When $\lambda_k(\bm{e}_s)=on$, there will be PBI-rep $\bm{\mathcal{I}}(\bm{e}_s)$ attached to $\bm{e}_s$ and the related neighborhood $\Psi_k(\bm{\mathcal{I}}(\bm{e}_s))$ must be nonzero, since it is on the surface of mesh $P_k$:
\begin{equation}
\label{eq:eq1}
\lambda_k(\bm{e}_s)=on,\ \ \  \mbox{iff. }\Psi_k(\bm{\mathcal{I}}(\bm{e}_s))\neq 0.
\end{equation}
So all we need is to iterate over all nonzero components of $\bm{\Psi}(\bm{\mathcal{I}}(\bm{e}_s))$ and change corresponding indicators to $on$.

\vspace{0.5em}
\noindent\textbf{$\bm{\bm{e}_s\to s}$}~~~~From eq. \ref{eq:porder}, we know if $\lambda_k(\bm{e}_s) \neq on$, then $\lambda_k(\bm{e}_s)=\lambda_k(s)$. When $\lambda_k(\bm{e}_s) = on$, we figure out $\lambda_k(s)$ using $\Psi_k(\bm{\mathcal{I}}(\bm{e}_s))$. From eq. \ref{eq:eq1}, we know $\Psi_k(\bm{\mathcal{I}}(\bm{e}_s)) \neq 0$ and can be either a triangle face or an edge. If $\Psi_k(\bm{\mathcal{I}}(\bm{e}_s))$ is a face, we can build a trivial BSP \cite{thibault1987set} using the face on the space near $\bm{e}_s$. If $\Psi_k(\bm{\mathcal{I}}(\bm{e}_s))$ is an edge, we can still build BSP using the opposite two faces along $\bm{e}_s$. In both cases, the BSP can be used to compute $\lambda_k(s)$ if we can sample a point from $\bm{s}$ that near $\bm{e}_s$. Unfortunately, we cannot guarantee to find such a point with precise coordinates. Instead, we choose a point $\bm{v}_x$ from the half plane $\bm{p}_{s, sp}^+$, which is on the supporting plane of $\bm{s}$ and on the positive side of bounding plane $\bm{p}_{s, b}^{\bm{e}_s}$ (Fig. x). For polygons, there is at least one such $\bm{v}_x$ among the vertices of $\bm{s}$. We find that the BSP is so simple that indicators $\lambda_k$ of all points on $\bm{p}_{s, sp}^+$ are the same. Since $\bm{p}_{s, sp}^+ \cap s \neq \varnothing$, we can compute $\lambda_k(s)$ by figuring out $\lambda_k(\bm{v}_x)$. Also, if we assign each BSP splitting plane a normal by the normal of triangle it contains, we can use it to decide the orientation ($same$ or $oppo$) when $\bm{v}_x$ is on a splitting plane, which means $\lambda_k(s)=on$.

\vspace{0.5em}
\noindent\textbf{$\bm{\bm{v}_{\bm{e}}\to e}$}~~~~ While this is yet another propagation from low dimension to high dimension, there is no neighborhood information stored on vertex. If $\lambda_k(\bm{v}_{\bm{e}})=on$, we first check whether $\lambda_k(e) = on$ by (\ref{eq:eq1}). If not, we need to find where $\bm{v}_{\bm{e}}$ lies on the surface of $P_k$ and determine the neighborhood of $\bm{v}_{\bm{e}}$. In most cases, it is straightforward since coincident vertices among different intersection-free meshes are shared. Exceptions are discussion is in \S\ref{sec:esubroutine}. The neighborhood may be a vertex, an edge or a face. After computing the neighborhood, the same BSP-based classification method is used as in ${\bm{e}_s\to s}$ to determine the high dimensional $\lambda_k(e)$. We use the other vertex of $e$ to be the sample point $\bm{v}_x$ for the BSP classification.

@@

Given a face $t_i$ to be classified and an I-rep $\boldsymbol\Gamma$ of intersection on one of $t_i$'s edge, we denote the three corners it as $\bm{v}_{b*}^{0}(t_i), \bm{v}_{b*}^{1}(t_i), \bm{v}_b^2(t_i),$ where the subscript $*$ means the vertices are on the edge where $\boldsymbol\Gamma$ lies. We also denote the context component of $\boldsymbol\Gamma$ as $C_{P_h \backslash P_j}$, meaning $t_i$ is from primitive $P_j$ and intersects primitive $P_h$. From previous discussion (section \ref{sec:ir}){\color{red}{may be more than one}}, we know there are two types of intersection context---face neighborhood and edge neighborhood. In both conditions, the one or two faces involving in the context split the space into two in and out space by its (their) orientation(s). We will see that under such division (denoted as $\boldsymbol{D}(C_{P_h \backslash P_j})$), we can compute $\lambda(t_i, P_j)$ exactly.
